---
title: "Machine Learning Course Project"
output: html_document
---

#Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement—a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


#Loading libraries
```{r}
library('caret')
library('doParallel')
```



#Enable multi-core processing

```{r}
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```

#Submission function(from Course project assignment)
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```

#Loading data
```{r}
data<-read.csv('pml-training.CSV')
test_data<-read.csv('pml-testing.CSV')
```

#Subsetting of the significant variables.
```{r}
useful_col<-c('roll_belt','pitch_belt','yaw_belt','total_accel_belt','gyros_belt_x','gyros_belt_y','gyros_belt_z','accel_belt_x','accel_belt_y','accel_belt_z','magnet_belt_x','magnet_belt_y','magnet_belt_z','roll_arm','pitch_arm','yaw_arm','total_accel_arm','gyros_arm_x','gyros_arm_y','gyros_arm_z','accel_arm_x','accel_arm_y','accel_arm_z','magnet_arm_x','magnet_arm_y','magnet_arm_z','roll_dumbbell','pitch_dumbbell','yaw_dumbbell','total_accel_dumbbell','gyros_dumbbell_x','gyros_dumbbell_y','gyros_dumbbell_z','accel_dumbbell_x','accel_dumbbell_z','magnet_dumbbell_x','magnet_dumbbell_y','magnet_dumbbell_z','roll_forearm','pitch_forearm','yaw_forearm','total_accel_forearm','gyros_forearm_x','gyros_forearm_y','gyros_forearm_z','accel_forearm_x','accel_forearm_y','accel_forearm_z','magnet_forearm_x','magnet_forearm_y','magnet_forearm_z','classe')

useful_col1<-c('roll_belt','pitch_belt','yaw_belt','total_accel_belt','gyros_belt_x','gyros_belt_y','gyros_belt_z','accel_belt_x','accel_belt_y','accel_belt_z','magnet_belt_x','magnet_belt_y','magnet_belt_z','roll_arm','pitch_arm','yaw_arm','total_accel_arm','gyros_arm_x','gyros_arm_y','gyros_arm_z','accel_arm_x','accel_arm_y','accel_arm_z','magnet_arm_x','magnet_arm_y','magnet_arm_z','roll_dumbbell','pitch_dumbbell','yaw_dumbbell','total_accel_dumbbell','gyros_dumbbell_x','gyros_dumbbell_y','gyros_dumbbell_z','accel_dumbbell_x','accel_dumbbell_z','magnet_dumbbell_x','magnet_dumbbell_y','magnet_dumbbell_z','roll_forearm','pitch_forearm','yaw_forearm','total_accel_forearm','gyros_forearm_x','gyros_forearm_y','gyros_forearm_z','accel_forearm_x','accel_forearm_y','accel_forearm_z','magnet_forearm_x','magnet_forearm_y','magnet_forearm_z')


data1<-subset(data,select=useful_col)
test_data1<-subset(test_data,select=useful_col1)

```

#Creating data partiton
```{r}
intrain<-createDataPartition(data1$classe,p=0.75,list=F)
training<-data1[intrain,]
testing<-data1[-intrain,]
```


#Training the model using preprocess
Preprocess was the centering and scaling. Type of the model was selected "gbm", Stochastic Gradient Boosting as a powerful and general method. 
```{r}
modelfit<-train(classe~.,data=training,method='gbm',preProcess=c("center", "scale"))
modelfit
confusionMatrix(testing$classe,predict(modelfit,testing))

```
The confusion matrix demonstrates that the model is suitable for prediction.


#Making the submission
```{r}
submission<-predict(modelfit, test_data1)
pml_write_files(submission)
submission
```

